Water discharge quantity has a variable impact on sediment transport capacity in subglacial channels

### 2023-09-21 Delaneyetal2023b.tex
getting ready to pass around
Updates from Overleaf
small changes to be passed around

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

@@

Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both \alpine{} and \icesheet{}.
This occurs as the subglacial values are generally larger than the subaerial values, thus more efficient at transporting sediment.
In the subglacial channel, peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
Conversely, in the subaerial channel, peaks in these sediment transport variables occur when water discharge is highest.
The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge.
Here, sediment could be deposited in the proglacial area as the water depressurizes, then remobilized when sediment transport capacity peaks in the subaerial system.

@@

Each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
Conversely in the subglacial channel,  water velocity, shear stress, and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two  outputs as well (Figure~\ref{fig:Qw_vari} a--f).
For instance, very high  water velocity values and shear stresses can occur across minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning much of the water discharge range (Figure~\ref{fig:Qw_vari} d--f).

@@

The second application examines the impact of a variety of channel shapes and friction factors on the relationship between water discharge and sediment transport capacity, in both systems.
Across the range of parameter values examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
However, in \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
Conversely, in \icesheet{}, with lower water discharge variability, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.

@@

The rapid increases and decreases in subglacial sediment transport capacity may compound the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
By applying the discharge - width scaling in Equation~\ref{eq:wcf} to Equation~\ref{eq:v}, by setting $S\, \propto \,Q^{\frac{1}{2}}$, the subaerial channel's width-integrated shear stress response is $\tau_{tf}\, \propto\,  Q$. 
For sufficiently short time periods,  channel width and wetted area are fixed in the subglacial system so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$, with a larger exponent on $Q$ than in the subaerial case following \citeA{alley1997}.
In this case, water discharge fluctuates at a shorter timescale than the subglacial channel evolution \cite{gimbert2016}.
Note also that the subglacial width integrated shear stress response to water discharge will be greater than the subaerial's provided that  Equation~\ref{eq:wcf}'s exponent is below one.
However, when subglacial channel size lies in equilibrium with water discharge, the  width-integrated shear stress response is $\tau_{tg}\, \propto \, Q^{\frac{4}{5}}$, similar to subaerial channels.
It follows that, the stronger correlation between hydraulics and water discharge in \icesheet{} could result from being closer to this equilibrium between variations in the channels wetter area and water discharge compared to \alpine{}.
This state could result from the reduced diurnal discharge amplitude in \icesheet{} (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}).
In both cases, hydraulics could trend toward this equilibrium regime when average discharge over $1-5$ days when variability in subglacial channels  approaches that of subaerial channels (Figure~\ref{fig:multi_run}).

@@

SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
We thank G. King for commenting on a previous version of this manuscript.

@@

\end{table}

@@

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.7\linewidth]{Fig3_hr.png}
    \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{hr} aggregation.}
    \label{fig:model_outs_1hr}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.7\linewidth]{Fig3_12hr.png}
    \caption{As Figure \ref{fig:model_outs}, with $12$ \,\unit{hr} aggregation.}
    \label{fig:model_outs_12hr}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.7\linewidth]{Fig3_1day.png}
    \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{day} aggregation.}
    \label{fig:model_outs_1day}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.7\linewidth]{Fig3_5day.png}
    \caption{As Figure \ref{fig:model_outs}, with $5$ \,\unit{day} aggregation.}
    \label{fig:model_outs_5day}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.7\linewidth]{Fig3_10day.png}
    \caption{As Figure \ref{fig:model_outs}, with $10$ \,\unit{day} aggregation.}
    \label{fig:model_outs_10day}
  \end{figure}
\end{center}
\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.7\linewidth]{Fig3_15day.png}
    \caption{As Figure \ref{fig:model_outs}, with $15$ \,\unit{day} aggregation.}
    \label{fig:model_outs_15day}
  \end{figure}
\end{center}

\section{Mauro's bit about shear stress}
% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options. This is common when trying to add a PDF figure without pdflatex.
% \begin{figure}
% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
%\caption{caption}
%\label{pdffiguresample}
%\end{figure}
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendices go here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}
\end{document}

>>>>

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage{placeins}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

@@

Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both \alpine{} and \icesheet{}.
This occurs as the subglacial values are generally larger than the subaerial values, thus more efficient at transporting sediment.
In the subglacial channel, peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
Alternatively, in the subaerial channel, peaks in these sediment transport variables occur when water discharge is highest.
The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge.
Here, sediment could be deposited in the proglacial area as the water depressurizes, then remobilized when sediment transport capacity peaks in the subaerial system.

@@

Each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).
In the subglacial channel, however,  water velocity, shear stress, and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two  outputs as well (Figure~\ref{fig:Qw_vari} a--f).
For instance, very high  water velocity values and shear stresses can occur across minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
In \icesheet, mean values of water velocity can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning much of the water discharge range (Figure~\ref{fig:Qw_vari} d--f).

@@

The second application examines the impact of a variety of channel shapes and friction factors on the relationship between water discharge and sediment transport capacity, in both systems.
Across the range of parameter values examined, variability in water velocity and shear stress remains higher in the subglacial system compared to the subaerial one.
However, in \alpine{}, the variability in width-integrated shear stress is comparable between the two systems, due to the variable water discharge (Figure~\ref{fig:multi_run}).
 However, in \icesheet{}, with lower water discharge variability, the variability in width-integrated shear stress remains higher in the subglacial system across the time aggregation periods.

@@

The rapid increases and decreases in subglacial sediment transport capacity may compound the variability in sediment transport already present in subaerial systems \cite<Figure~\ref{fig:Qw_vari}; >{williams1989,jerolmack2010}.
By applying the discharge - width scaling in Equation~\ref{eq:wcf} to Equation~\ref{eq:v}, by setting $S\, \propto \,Q^{\frac{1}{2}}$, the subaerial channel's width-integrated shear stress response is $\tau_{tf}\, \propto\,  Q$ (see \ref{sect:scaling}). 
For sufficiently short time periods,  channel width and wetted area are fixed in the subglacial system so that the response of width-integrated shear stress is $ \tau_{tg}\, \propto\,  Q^{2}$, with a larger exponent on $Q$ than in the subaerial case following \citeA{alley1997}.
In this case, water discharge fluctuates at a shorter timescale than the subglacial channel evolution \cite{gimbert2016}.
Note also that the subglacial width integrated shear stress response to water discharge will be greater than the subaerial's provided that  Equation~\ref{eq:wcf}'s exponent is below one.
However, when subglacial channel size lies in equilibrium with water discharge, the  width-integrated shear stress response is $\tau_{tg}\, \propto \, Q^{\frac{4}{5}}$, similar to subaerial channels.
It follows that the stronger correlation between hydraulics and water discharge in \icesheet{} could result from being closer to this equilibrium between variations in the channels' wetter area and water discharge compared to \alpine{}.
This state could result from the reduced diurnal discharge amplitude in \icesheet{} (Figures~\ref{fig:model_outs} and ~\ref{fig:Qw_vari}).
In both cases, hydraulics could trend toward this equilibrium regime when average discharge over $1-5$ days when variability in subglacial channels  approaches that of subaerial channels (Figure~\ref{fig:multi_run}).

@@

SNSF Project No. PZ00P2\_202024 provided  funding for I. Delaney.
G. King provided useful comments on a previous version of this manuscript.

@@

\end{table}
\FloatBarrier

@@

\begin{center}
  \begin{figure}[!h]
    \includegraphics[width=0.7\linewidth]{Fig3_hr.png}
    \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{hr} aggregation.}
    \label{fig:model_outs_1hr}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[!h]
    \includegraphics[width=0.7\linewidth]{Fig3_12hr.png}
    \caption{As Figure \ref{fig:model_outs}, with $12$ \,\unit{hr} aggregation.}
    \label{fig:model_outs_12hr}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[!h]
    \includegraphics[width=0.7\linewidth]{Fig3_1day.png}
    \caption{As Figure \ref{fig:model_outs}, with $1$ \,\unit{day} aggregation.}
    \label{fig:model_outs_1day}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[!h]
    \includegraphics[width=0.7\linewidth]{Fig3_5day.png}
    \caption{As Figure \ref{fig:model_outs}, with $5$ \,\unit{day} aggregation.}
    \label{fig:model_outs_5day}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[!h]
    \includegraphics[width=0.7\linewidth]{Fig3_10day.png}
    \caption{As Figure \ref{fig:model_outs}, with $10$ \,\unit{day} aggregation.}
    \label{fig:model_outs_10day}
  \end{figure}
\end{center}

\begin{center}
  \begin{figure}[!h]
    \includegraphics[width=0.7\linewidth]{Fig3_15day.png}
    \caption{As Figure \ref{fig:model_outs}, with $15$ \,\unit{day} aggregation.}
    \label{fig:model_outs_15day}
  \end{figure}
\end{center}
\FloatBarrier

\newpage
\section{Mauro's bit about shear stress}
\label{sect:scaling}
% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options. This is common when trying to add a PDF figure without pdflatex.
% \begin{figure}
% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
%\caption{caption}
%\label{pdffiguresample}
%\end{figure}
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendices go here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}
\end{document}
